### 2019-03-01 Seneca.tex
mz - michael pass
Rich makes a pass

Based on the specified machine learning application, Seneca
automatically builds and deploys an AWS Lambda application by
launching a Docker container that mirrors the AWS Lambda execution
environment, checks and installs the machine learning application and
any libraries it requires, compresses the application and uploads it
to S3 (a
work-around for the 10MB AWS Lambda function size restriction).
Seneca constructs an AWS Lambda function from a template that, when
executed, will download the dataset and split it into a training and
testing set, and construct, test, and evaluate a model using the
application and a set of hyperparameter values passed in by Seneca as
arguments.  Users can specify the train/test split ratio that should
be used by Seneca; the default is 80\%/20\%.  The function returns a
testing score.  Upon completion of this process, the container deploys
the function to AWS Lambda using the AWS Command Line
Interface~\cite{ref:awscli}
%\footnote{\url{https://aws.amazon.com/cli/}} 
and the developers credentials.

@@

We assume that the dataset supplied to Seneca by the user is representative of 
datasets on which the 
resulting model will be used.  In addition, we use prediction error as the 
score (i.e., mean squared error for regression and accuracy percentage for classification) 
instead of $R^2$, which describes explanatory power, to avoid overfitting.
As part of future work, we are considering using multiple datasets and a ranges
of hyperparameter values to preclude the need for users to specify
them and to consider a wider range of values.

>>>>

Based on the specified machine learning application, Seneca
automatically builds and deploys an AWS Lambda application by
launching a Docker container that mirrors the AWS Lambda execution
environment, checks and installs the machine learning application and
any libraries it requires, compresses the application and uploads it
to S3 (a
work-around for the 10MB AWS Lambda function size restriction).
Seneca constructs an AWS Lambda function from a template that, when
executed, will download the dataset and split it into a training and
testing set, and construct, test, and evaluate a model using the
application and a set of hyperparameter values passed in by Seneca as
arguments.  Users can specify the train/test split ratio that should
be used by Seneca; the default is 80\%/20\% for classification tasks.  The function returns a
testing score.  Upon completion of this process, the container deploys
the function to AWS Lambda using the AWS Command Line
Interface~\cite{ref:awscli}
%\footnote{\url{https://aws.amazon.com/cli/}} 
and the developers credentials.

@@

We assume that the dataset supplied to Seneca by the user is representative of 
datasets on which the 
resulting model will be used.  
%In addition, we use prediction error as the 
%score (i.e., mean squared error for regression and accuracy percentage for classification) 
%instead of $R^2$, which describes explanatory power, to avoid overfitting.
As part of future work, we are considering using multiple datasets and a ranges
of hyperparameter values to preclude the need for users to specify
them and to consider a wider range of values.

### 2019-03-01 intro.tex
Rich makes a pass

Seneca implements, packages, and deploys 
machine learning applications as stateless functions to AWS Lambda.
It then orchestrates exhaustive evaluation of specified hyperparameter settings
to identify the best performing model (for a given dataset) by
comparing error and accuracy across models.  We consider
prediction accuracy (as opposed to explanatory power) as the
scoring metric (mean squared error for regression 
and accuracy percentage for classification), to avoid overfitting.
Users present Seneca with their application, a range of values for 
each hyperparameter (or the default can be used), and a representative dataset.
Seneca produces, tests, and evaluates models for all combinations 
of hyperparameters and returns to the user
the set of parameters (or the model itself) that produces
the best cross-validation score.
Users can employ this model for other datasets
(with Seneca if desired) without retraining the model to amortize the cost of Seneca further.

We deploy Seneca on AWS Lambda and evaluate its tuning performance, cost,
and memory use for five machine learning applications and datasets. We 
find that Seneca is fast, inexpensive, and effective for model 
construction and comparison. Seneca is also able to identify
automatically the
best memory configuration for each application, further lowering its
cost
by 10-35\%. Relative to execution in AWS EC2, Seneca enables average speedups of 294x for each additional dollar 
spent for the applications and datasets we consider.
We intend to make Seneca, its applications, and their datasets publicly available
when/if this paper is accepted for publication.  
We next overview our design and implementation of Seneca and then
present our empirical methodology and results.

>>>>

Seneca implements, packages, and deploys 
machine learning applications as stateless functions to AWS Lambda.
It then orchestrates exhaustive evaluation of specified hyperparameter settings
to identify the best performing model (for a given dataset) by
comparing error and accuracy across models.  We consider
prediction accuracy (as opposed to explanatory power) as the
scoring metric (mean squared prediction error for regression 
and accuracy percentage for classification).
Users present Seneca with their application, a range of values for 
each hyperparameter (or the default can be used), and a representative dataset.
Seneca produces, tests, and evaluates models for all combinations 
of hyperparameters and returns to the user
the set of parameters (or the model itself) that produces
the best cross-validation score.
Users can employ this model for other datasets
(with Seneca if desired) without retraining the model to amortize the cost of Seneca further.

We deploy Seneca on AWS Lambda and evaluate its tuning performance, cost,
and memory use for five machine learning applications and datasets. We 
find that Seneca is fast, inexpensive, and effective for model 
construction and comparison. Seneca is also able to identify
automatically the
best memory configuration for each application, further lowering its
cost
by 10-35\%. Relative to execution in AWS EC2, we observe a benefit/cost
ratio (computed as speedup/(dollar cost) cost of 294 on the average
in the experiments we consider.
We intend to make Seneca, its applications, and their datasets publicly available
when/if this paper is accepted for publication.  
We next overview our design and implementation of Seneca and then
present our empirical methodology and results.

### 2019-03-01 results.tex
Rich makes a pass

\begin{table}[t]
\centering
\input{tables/application-tab}
\caption{Machine learning applications used 
to evaluate Seneca. 
%paper submission is blind (so such references must be omitted so as not to reveal our identities):
%Code base is available at project repository~\cite{ref:seneca}.
\label{tab:bmarks}}
\vspace{-0.1in}
\end{table}

@@

To evaluate Seneca, we measure model output quality, execution time, memory use, and monetary cost.  For output quality (prediction accuracy) our metrics are mean squared error (MSE) for regression and percentage accuracy for classification as described above. We score model prediction accuracy and not explanatory power ($R^2$) to avoid overfitting.

@@

\begin{figure}[t] \centering 
\vspace{-0.5in}
\includegraphics[scale=0.36]{box_plot_mse}
\caption{Box plot of MSE from the three regression applications across the 
hyperparameter tuning search space. The red notch shows the MSE from the default settings. 
The colored diamonds are outliers beyond two interquartile ranges. 
Seneca selects the points indicated by blue triangle.
Lower MSE values are better. 
\label{fig:box_plot_mse}}
\vspace{-0.2in}
\end{figure}

We first evaluate the quality of the output generated by each
ML applications when Seneca determines the hyperparameter settings. We
first show the results for the regression applications in
Table~\ref{tab:mse}.
The first row of data is the number of hyperparameter 
configurations that Seneca considers for each.
The last three rows show the MSE for the default, worst, and best performing (Seneca's recommendation)
hyperparameter configuation (lower is better).
Seneca reduces MSE by 22.56\%, 15.94\%, and 44.88\%, for Prophet, Multi-Regression, and XGBoost,
respectively, for the datasets and training methodologies that we consider.
Compared to the worst case, Seneca reduces MSE by 82.62\%, 78.01\%, and 99.28\%, respectively.

@@

\begin{figure}[t] \centering 
\vspace{-0.1in}
\includegraphics[scale=0.36]{box_plot_Accuracy_20}
\vspace{-0.4in}
\caption{Box plot of accuracy reported for three classification applications across 
the hyperparameter search space. The red notch indicates the accuracy that results
from default hyperparameter values.
The diamonds are outliers beyond two interquartile ranges. 
Seneca selects the points indicated by blue triangle. 
Higher accuracy is better. 
\label{fig:box_plot_accuracy}}
\vspace{-0.2in}
\end{figure}

@@

\begin{table}
\centering
\input{tables/yield-tab-corrected.tex}
\caption{Seneca VS EC2 cost analysis. Execution time (mins) and cost (cents) for
executing the applications serially in EC2 (t2.medium).  
Rows 3--4 show yield -- the additional speedup that
Seneca can achieve for each additional dollar spent for these applications.
Yield for SVC is 0 (infinite) because Seneca costs less than EC2 in this case.
Ideal yield shows the yield when we execute the applications in parallel (assuming
2x perfect parallelism).
\label{tab:yield}}
\vspace{-0.1in}
\end{table}

Finally, we compare the cost of Seneca to AWS Elastic Compute Cloud (EC2) use. 
We measure the execution time of Seneca using the
least expensive EC2 instance type in which the applications will run
(t2.medium, which has 2 multi-tenant cores and 4GB of memory). 
Note that EC2 instances are charged for by the hour; Lambda charges 
are only imposed when functions execute. We execute the applications serially
using the instance.  For this setting, Seneca enables a speedup over EC2 
of 3.72x -- 12.38x (6.51x on average). Doing so, however, imposes an additional 
cost of \$0.01--\$0.15 over EC2 for all but SVC. 
Seneca costs \$0.03 less than EC2 for SVC (because SVC runs in significantly less than
the next hour boundary).

To further understand the relationship between Seneca speedup and cost when Seneca
is more expensive (but faster) than using EC2, we define \texttt{yield} as 
$Y=\frac{T_{ec}}{T_{sc}}/({C_{sc} - C_{ec}}) \,|\ if \,C_{sc} > C_{ec}$ where $T_{ec}$ and $T_{sc}$ are the execution time, $C_{ec}$ and $C_{sc}$ are the total cost of EC2 instance 
and Seneca, respectively.
For applications for which Seneca is cheaper (e.g. SVC), we report \texttt{yield} 
as $0.00$.  This metric reveals the amount of speed up that Senence can achieve 
for each additional dollar spent.  To understand the yield if we were to 
parallelize the EC2 deployment (perhaps a more ``fair'' comparison), 
we also estimate yield for perfect parallelism (2x in our case for 
the t2.medium).

We present results for this yield metric in Table~\ref{tab:yield} for each of the applications.
Rows 1 and 2 show the average execution time (in minutes) and cost (in cents) 
from using EC2 for each.
Rows 3 and 4 show the Seneca yield (speedup/\$).
Row 3 shows yield for serialized execution in EC2 (t2.medium
instance) and row 4 shows estimated yield if we were to achieve perfect parallelism
(i.e. 2x) using the EC2 instance. On average across the four applications for which
EC2 is cheaper, Seneca achieves yield of 294 (assuming perfect parallelism in EC2). 
That is, Seneca is able to provide a speedup
of 294x on average, for each additional dollar spent for these applications.
We plan to compare Seneca cost and performance 
to other EC2 instances and the AWS Elastic Container Service
as part of future work.

>>>>

\begin{table}[t]
\centering
\input{tables/application-tab}
\caption{Machine learning applications used 
to evaluate Seneca. 
%paper submission is blind (so such references must be omitted so as not to reveal our identities):
%Code base is available at project repository~\cite{ref:seneca}.
\label{tab:bmarks}}
\vspace{-0.2in}
\end{table}

@@

To evaluate Seneca, we measure model output quality, execution time, memory
use, and monetary cost.  For output quality (prediction accuracy) our metrics
are mean squared prediction error (MSE) for regression and percentage accuracy for
classification as described above. 
%We score model prediction accuracy and not
%explanatory power ($R^2$) to avoid overfitting.

@@

\begin{figure}[t] \centering 
\vspace{-0.5in}
\includegraphics[scale=0.36]{box_plot_mse}
\vspace{-0.4in}
\caption{Box plot of MSE from the three regression applications across the 
hyperparameter tuning search space. The red notch shows the MSE from the default settings. 
The colored diamonds are outliers beyond two interquartile ranges. 
Seneca selects the points indicated by blue triangle.
Lower MSE values are better. 
\label{fig:box_plot_mse}}
\vspace{-0.1in}
\end{figure}

We first evaluate the quality of the output generated by each
ML applications when Seneca determines the hyperparameter settings. We
first show the results for the regression applications in
Table~\ref{tab:mse}.
The first row of data is the number of hyperparameter 
configurations that Seneca considers for each.
The last three rows show the MSE for the default, worst, and best performing (Seneca's recommendation)
hyperparameter configuration (lower is better).
Seneca reduces MSE by 22.56\%, 15.94\%, and 44.88\%, for Prophet, Multi-Regression, and XGBoost,
respectively, for the datasets and training methodologies that we consider.
Compared to the worst case, Seneca reduces MSE by 82.62\%, 78.01\%, and 99.28\%, respectively.

@@

\begin{figure}[t] \centering 
\vspace{-0.3in}
\includegraphics[scale=0.36]{box_plot_Accuracy_20}
\vspace{-0.4in}
\caption{Box plot of accuracy reported for three classification applications across 
the hyperparameter search space. The red notch indicates the accuracy that results
from default hyperparameter values.
The diamonds are outliers beyond two interquartile ranges. 
Seneca selects the points indicated by blue triangle. 
Higher accuracy is better. 
\label{fig:box_plot_accuracy}}
\vspace{-0.1in}
\end{figure}

@@

\begin{table}
\centering
\input{tables/yield-tab-corrected.tex}
\caption{Seneca VS EC2 cost analysis. Execution time (mins) and cost (dollars) for
executing the applications serially in EC2 (t2.medium).  
Rows 3--4 show yield -- the additional speedup that
Seneca can achieve for each additional dollar spent for these applications.
Yield for SVC is 0 (infinite) because Seneca costs less than EC2 in this case.
Ideal yield shows the yield when we execute the applications in parallel (assuming
2x perfect parallelism).
\label{tab:yield}}
\vspace{-0.1in}
\end{table}

Finally, we compare the cost of Seneca to the cost of using AWS Elastic Compute Cloud (EC2). 
We measure the execution time of Seneca using the
least expensive EC2 instance type in which the applications will run
(t2.medium, which has 2 multi-tenant cores and 4GB of memory). 
Note that EC2 instances are charged for by the hour; Lambda charges 
are only imposed when functions execute, but Lambda \textit{may} execute
the functions concurrently. 
For this comparison, Seneca (because Lambda's automatic exploitation of
concurrency) enables a speedup over EC2 (where we execute the parameter sweep
sequentially) 
of 3.72x -- 12.38x (6.51x on average). This speedup comes at an additional 
cost of \$0.01--\$0.15 over EC2 for all but SVC (which executes for
significantly less than an hour). 
%Seneca costs \$0.03 less than EC2 for SVC (because SVC runs in significantly less than
%the next hour boundary).

To further understand the relationship between Seneca speedup and cost when Seneca
is more expensive (but faster) than using EC2, we define \texttt{yield} as 
$Y=\frac{T_{ec}}{T_{sc}}/({C_{sc} - C_{ec}}) \,|\ if \,C_{sc} > C_{ec}$ where $T_{ec}$ and $T_{sc}$ are the execution time, $C_{ec}$ and $C_{sc}$ are the total cost of EC2 instance 
and Seneca, respectively.
For applications for which Seneca is cheaper (e.g. SVC), we report \texttt{yield} 
as $0.00$ since there is no positive benefit/cost ratio.  This metric captures
the amount of speed up that Seneca can achieve 
for each additional dollar spent.  To make the comparison ``fair''
we also explore yield for theoretically perfect parallelism in EC2 using 2
cores (e.g. in a t2.medium).

We present results for this yield metric in Table~\ref{tab:yield} for each of the applications.
Rows 1 and 2 show the average execution time (in minutes) and cost (in dollars) 
from using EC2 for each.
Rows 3 and 4 show the Seneca yield (speedup/\$).
Row 3 shows yield for serialized execution in EC2 (t2.medium
instance) and row 4 shows estimated yield if we were to achieve perfect parallelism
(i.e. 2x) using the EC2 instance. On average across the four applications for which
EC2 is cheaper, Seneca achieves yield of 294 (assuming perfect parallelism in EC2). 
That is, Seneca is able to provide a speedup
of 294x on average, for each additional dollar spent for these applications.
It is tempting to attribute this effect to the EC2 full-hour charge for short
running applications, but XGBoost runs for almost 6 hours and the yield is
still 41.86 assuming best-case EC2 execution.  We plan to study this
benefit/cost ratio for Seneca and Lambda applications 
as part of future work.
